---
title: "Grandparent_trios"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Grandparent_trios}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gRandma)
```


This will walk through an example of estimating error rates for assigning grandparent-grandchild trios 
and inferring grandparent-grandchild trios.


## Setup

First, we make a gmaData object. For details, see the "Load_in_data_and_gmaData_structure" vignette.

```{r}
# removing loci with no variation in the subset of pops we are going to use
tempData <- data_mh_snp[data_mh_snp$Pop %in% c("Pop_2", "Pop_4"),]
tRem <- c()
for(i in seq(3, ncol(tempData) - 1, 2)) if(length(na.omit(unique(c(tempData[[i]], tempData[[i+1]])))) < 2) tRem <- c(tRem, i)
tempData <- tempData[,-c(tRem, tRem + 1)]
# creating a gmaData object
gData_1 <- createGmaInput(baseline = tempData[tempData$Pop == "Pop_2",], 
								  mixture = tempData[tempData$Pop == "Pop_4",-1], 
								  perAlleleError = .005, dropoutProb = .005,
										  markerType = "microhaps", alleleDistFunc = (function(x) return(1)))
```

## Error rate estimation

### False negative error rates

We are first going to estimate false negative error rates. This is the probability that a true 
grandparent-grandchild trio fails to assign. Here we are testing critical values ($c$ in the manuscript) 
of $0, 1, 2, ..., 25$. 

```{r}
falseNegative <- falseGrandma(gData_1, relationship = "ssGP", 
										llrToTest = seq(0, 25, 1), N = 10000,
										seed = 7, errorType = "falseNegative")
falseNegative[[1]][1:10,]
```

We now have an estimate of false negative error rates for each critical value for our one baseline population.

### False positive error rates

Let's first consider false positive error rates for unrelated trios. We 
can estimate these rates by importance sampling, here running 10,000 iterations (you should run more), and testing 
critical values from 0 to 9.

```{r}
falsePos1 <- falseGrandma(gData_1, relationship = "ssGP", 
										llrToTest = seq(0, 9, 1), N = 10000,
										seed = 7, errorType = "Unrel", method = "IS")
falsePos1[[1]]
```

We now have false positive estimates. When false positive error rates are computed by importance sampling, 
it is computationally almost free to also calculate false negative rates, so they are also reported. We can 
see that this panel is not very powerful, with our false positive error 
rate on the order of $10^{-6}$ with the false negative rate approaching $0.05$. 

We can also calculate these error rates by stratified sampling. We will run 1,000 iterations in each stratum. This 
is too few, but we will use it here because it is quick to build the vignette and because it will allow us to demonstrate 
the main drawback of the stratified sampling approach. 
If you wanted, you could use a different number of iterations in each stratum to target computational effort. The 
input, `rep(1000, 10)`, tells it to run 1000 iterations in the first 10 strata (number of MIs from 0 - 9). We know from 
the false negative output above that it is only going to consider 0 - 4 MIs, so the function will stop after 4. Giving a number of 
iterations for extra strata is ok, but it will throw an error if you give too few. 

```{r}
falsePos2 <- falseGrandma(gData_1, relationship = "ssGP", 
										llrToTest = seq(0, 9, 1), itersPerMI = rep(1000, 10),
										seed = 7, errorType = "Unrel", method = "strat")
```

The output for stratified sampling is a list of 2. The first item has the overall false positive estimates:

```{r}
falsePos2[[1]]
```

The second item has the estimates for each stratum:

```{r}
falsePos2[[2]][1:10,]
```

The columns are:

* `count_MI`: the number of observed Mendelian incompatibilities (MIs) that stratum corresponds to
* `prob_count_MI`: the probability a trio of the defined relationship (Unrelated in this case)
  has that number of MIs observed
* `falsePosUnrel`: the false positive error rate estimated for that stratum. To get the number of
  observed false positives during the simulation, multiply this number by the number of iterations 
  run for that stratum (in this case, it was 1000)
* `falsePosUnrelSD`: the estimated standard deviation for this stratum

We can see right away that even for our lowest critical value of 0, the 4 MI stratum had 0 false 
positives observed and the 3 MI stratum only had 1. So, obviously we need to increase the number 
of iterations run for these strata to get a more reliable estimate. 

Now how about trios of other relationships? The process and outputs are the same, we just change the 
`errorType` argument to `falseGrandma`. For considering grandparent-grandchild trios, valid options 
for `errorType` are: `"falseNegative", "Unrel", "True_GAunt", "True_Unrel", "True_HGAunt", "True_GpCous", "GAunt_Unrel",
 "HGAunt_Unrel", "GpCous_Unrel", "GAunt", "GAunt_HGAunt", "Gaunt_GpCous", "HGAunt",
 "HGAunt_GpCous"`, and `"GpCous"`. For example:
 
```{r, eval=FALSE}
# not run
# importance sampling
falsePos_gAuntIS <- falseGrandma(gData_1, relationship = "ssGP", 
             llrToTest = seq(0, 9, 1), N = 1000000,
             seed = 7, errorType = "GAunt", method = "IS")
# stratified sampling
falsePos_gAuntStrat <- falseGrandma(gData_1, relationship = "ssGP", 
             llrToTest = seq(0, 9, 1), itersPerMI = rep(100000, 10),
             seed = 7, errorType = "GAunt", method = "strat")
```


## Assigning grandparents

Now we've estimated the error rates for this genetic panel and population, and for the sake of 
the example, let's say we've decided on a critical value of 8. Let's now look for grandparent-grandchild trios. 
We've already set up the gmaData object to have baseline individuals (potential grandparents) and 
mixture individuals (potential grandchildren).

Consider that by looking at metadata from the hatchery (sex, date of spawning, cross records, etc.) and/or 
a genetic sex marker, you determine that these are the possible crosses that could have been made:

```{r}
potentialCrosses <- data.frame(Pop = "Pop_2",
										 gp1 = gData_1$baseline[[2]][seq(1, 123, 2)],
										 gp2 = gData_1$baseline[[2]][seq(2, 124, 2)]
										 )
```


This will run the assignments and consider all combinations of grandparents listed in
`potentialCrosses`. If you want to consider all possible combinations of grandparents within each given 
baseline population, just omit the `crossRecords` argument.

```{r}
assignments <- inferGrandma(gData_1, relationship = "ssGP", minLLR = 8, crossRecords = potentialCrosses)
assignments
```

And we have one grandparent-grandchild trio assigned. 
